Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1].}

### 2019-02-27 paper/paper.tex
Fixes

Exploiting multiple cores to speed up constraint programming solvers remains an active area of
research, with no universally perfect solution being available. Four of the more common approaches
are based upon decompositions
\cite{DBLP:journals/corr/abs-1008-4328,DBLP:journals/jair/MalapertRR16}, work stealing
\cite{DBLP:conf/cp/MichelSH07,DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018},
parallel discrepancy searches \cite{DBLP:conf/cp/MoisanGQ13,DBLP:conf/cpaior/MoisanQG14}, and
algorithm portfolios \cite{DBLP:conf/lion/LindauerHH15}. Decomposition approaches are unsuitable for
decision problems, or problems where we have good value-ordering heuristics, because the
decomposition interferes strongly with the shape of the search tree
\cite{DBLP:journals/jair/MalapertRR16}. Work stealing, traditionally, also interferes with
value-ordering \cite{DBLP:journals/topc/McCreeshP15}, although specially designed exceptions exist
\cite{DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018}. However,
these have very complicated implementations. Parallel discrepancy searches are aware of
value-ordering heuristics, but have other limitations: they struggles on search trees with heavy
filtering, and rely upon inner search tree nodes being orders of magnitude less expensive to process
than leaf nodes. Portfolios, meanwhile, typically rely upon running multiple models or heuristics
simultaneously, and selecting whichever finishes first, whereas here we have a known good model and
set of heuristics.

@@

\Cref{figure:parallel} also shows the results of these experiments, using five, ten, and twenty
hosts.  Because each host has two CPU sockets, the five host results use ten MPI ranks, each with
eighteen threads, and the ten and twenty host results use twenty and forty MPI ranks respectively.
The supercomputing service we use is not designed for huge numbers of very short problems, and so we
ran only instances whose sequential runtime was at least one second; for the sake of plotting
results, we treat skipped instances as taking one second. Due to the job launcher used, it is also
not possible to accurately measure ``total'' runtime including startup costs, and so instead we
report the runtime of the rank zero process---this figure is somewhat optimistic for very easy
instances. With these caveats in mind, when seeing how long a timeout is needed to solve any 14,370
instances, we get aggregate speedups of 57.1 from five hosts and 95.5 from ten hosts over a
sequential baseline; using twenty hosts is slightly slower, due to increased overheads. However,
when seeing how long a timeout is needed to solve 14,424 instances, distributed parallelism gives us
further aggregate speedups of 2.2, 3.5, and 7.9 from five, ten, and twenty hosts, over the single
machine parallel baseline.

>>>>

Exploiting multiple cores to speed up constraint programming solvers remains an active area of
research, with no universally perfect solution being available. Four of the more common approaches
are based upon decompositions
\cite{DBLP:journals/corr/abs-1008-4328,DBLP:journals/jair/MalapertRR16}, work-stealing
\cite{DBLP:conf/cp/MichelSH07,DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018},
parallel discrepancy searches \cite{DBLP:conf/cp/MoisanGQ13,DBLP:conf/cpaior/MoisanQG14}, and
algorithm portfolios \cite{DBLP:conf/lion/LindauerHH15}. Decomposition approaches are unsuitable for
decision problems, or problems where we have good value-ordering heuristics, because the
decomposition interferes strongly with the shape of the search tree
\cite{DBLP:journals/jair/MalapertRR16}. Work-stealing, traditionally, also interferes with
value-ordering \cite{DBLP:journals/topc/McCreeshP15}, although specially designed exceptions exist
\cite{DBLP:conf/cp/ChuSS09,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/HoffmannMNPRS018}. However,
these have very complicated implementations. Parallel discrepancy searches are aware of
value-ordering heuristics, but have other limitations: they struggles on search trees with heavy
filtering, and rely upon inner search tree nodes being orders of magnitude less expensive to process
than leaf nodes. Portfolios, meanwhile, typically rely upon running multiple models or heuristics
simultaneously, and selecting whichever finishes first, whereas here we have a known good model and
set of heuristics.

@@

\Cref{figure:parallel} also shows the results of these experiments, using five, ten, and twenty
hosts.  Because each host has two CPU sockets, the five host results use ten MPI ranks, each with
eighteen threads, and the ten and twenty host results use twenty and forty MPI ranks respectively.
The supercomputing service we use is not designed for huge numbers of very short problems, and so we
ran only instances whose sequential runtime was at least one second; for the sake of plotting
results, we treat skipped instances as taking one second. Due to the job launcher used, it is also
not possible to accurately measure ``total'' runtime including startup costs, and so instead we
report the runtime of the rank zero process---this figure is somewhat optimistic for very easy
instances. With these caveats in mind, when seeing how long a timeout is needed to solve any 14,370
instances, we get aggregate speedups of 57.1 from five hosts and 95.5 from ten hosts over a
sequential baseline; using twenty hosts is slightly slower, due to increased overheads. Finally, if
we look at harder instances, by allow a longer sequential timeout, we can solve 14,415 instances
sequentially with the last at 8,549 seconds; at this difficulty level, we achieve aggregate speedups
of 82.6, 103.2, and 144.9 from five, ten and twenty hosts.
